A Bayesian Perspective \\\large{Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest}

### 2023-02-13 manuscript/manuscript.qmd
grammer corrections

Three different informative priors were considered i)  a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`) ii) a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`) iii) a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`) distributions to represent all the combined available RCT data[@RN6759;@RN6751], only the ARREST[@RN6751] data  and only the PRAGUE[@RN6759] data, respectively

These different prior probabilities were updated using with the INCEPTION ITT data to create the posterior distributions displayed in Table 2. The probability for enhanced eCPR survival has increased to 80.4% with a skeptical prior to 84.9% with the enthusiastic prior and, as expected, the associated uncertainty has been reduced, as reflected by the narrower 95% CrI, with the additional data. The posterior probability of improved eCPR survival goes from 80% with a skeptical prior to 85% with the enthusiastic prior.

The probability that the eCPR survival improvements exceed a minimum 10% clinical threshold for improvement are `r round(skep_prob_1.1*100,1)`% and `r round(enth_prob_1.1*100,1)`% for the skeptical and enthusiastic priors, respectively. The corresponding ROPE probabilities are `r round(skep_eq*100,1)`%  and `r round(enth_eq*100,1)`% (Table 2). Posterior probabilities with the combined prior fell as expected between the results with the skeptical and enthusiastic priors. The graphical presentations of these results are shown in Figure 2.

INCEPTION did not report a per-protocol analysis. Such an analysis may be helpful in assessing treatment efficacy, unfortunately at the risk of an increased risk of bias by not respecting the ITT principle. From INCEPTION Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the cCPR group compared to 5 survivors from 46 patients receiving eCPR. With a vague prior, the OR of increased survival with eCPR is decreased compared to cCPR but with very wide CrI (OR 0.45, 95% CrI 0.15 - 1.35), limiting any definitive conclusions. The decreased eCPR success rates in the INCEPTION PP data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR. The probabilities of eCPR survival compared to cCPR using the per protocol data and incorporating the previous identified informative priors are shown in Table 3.

@@

This Bayesian analysis of the INCEPTION ITT data alone (with a vague prior) suggests the presence of a non-negligible probability of benefit with eCPR, in the range of 73%, compared to cCPR. The probability that this improvement in survival exceeds 10% is almost 66%. The consideration of pertinent prior information from two previous RCTs[@RN6751;@RN6759] in the field results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%. Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability. In contrast to the INCEPTION original conclusion of similar survival effects between the two treatments, this re-analysis suggests only a modest probability of approximately 15% that survival probabilities between the two techniques are within a reasonably acceptable equivalency range. Additional uncertainty about the benefits or harms of eCPR arise from the per protocol analysis, both with and without informative priors.

The INCEPTION researchers have addressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion. However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data, and prevents a full and comprehensive exploitation and updating of past knowledge. For trials such as INCEPTION that fail to meet statistical significance and are often incorrectly thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765]. Rather than dichotomizing results into statistical significance or not, with an obligatory loss of information, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.

Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to 
individual data from any of the trials limiting the possibility of examining specific subgroups. Also for the per protocol analysis, this data was only available for survival and not the primary INCEPTION outcome that included survival with intact neurological status. The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices. Given the existence of only 3 trials  estimating the very divergent baseline cCPR rates could not be reliably modeled. We choose not to perform a traditional Bayesian meta-analysis with
a semi-informative prior for the between-study variation  because our
goal was to illustrate how Bayesian principles could be informative
when applied to the analysis and interpretation of new trial data, both in the situations of no and considerable previous knowledge.

The "take home" message from this study is that standard statistical analyses resulting in a conclusion of "similar survival effects" may be overly simplified and potentially inaccurate. This Bayesian analysis demonstrates that any definitive conclusions regarding the superiority, inferiority, or equivalence of eCPR to cCPR are lacking. Rather the possibility of a clinically meaningful benefit, or less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties. The final dilemma confronting clinicians is that even Bayesian analyses of randomized trials provide only probability estimates for *average* treatment effect and not for the elusive *individual* treatment effect.

>>>>

Three different informative priors were considered i)  a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`) ii) a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`) iii) a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`) distributions to represent all the combined available RCT data[@RN6759;@RN6751], only the ARREST[@RN6751] data  and only the PRAGUE[@RN6759] data, respectively. These different prior probabilities were updated using with the INCEPTION ITT data to create the posterior distributions displayed in Table 2. The posteriorprobability for enhanced eCPR survival has increased to 80.4% with a skeptical prior, to 84.9% with the enthusiastic prior and, as expected, the associated uncertainty has been reduced, as reflected by the narrower 95% CrI, with the additional data.

The probability that the eCPR survival improvements exceed a minimum 10% clinical threshold for improvement are `r round(skep_prob_1.1*100,1)`% and `r round(enth_prob_1.1*100,1)`% for the skeptical and enthusiastic priors, respectively. The corresponding ROPE probabilities are `r round(skep_eq*100,1)`%  and `r round(enth_eq*100,1)`% (Table 2). Posterior probabilities with the combined prior were as expected between the results with the skeptical and enthusiastic priors. The graphical presentations of these results are shown in Figure 2.

INCEPTION did not report a per-protocol analysis. Such an analysis may be helpful in assessing treatment efficacy, unfortunately at the risk of an increased risk of bias by not respecting the ITT principle. From INCEPTION Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the cCPR group compared to 5 survivors from 46 patients receiving eCPR. With a vague prior, the OR of increased survival with eCPR compared to cCPR is decreased but with very wide CrI (OR 0.45, 95% CrI 0.15 - 1.35), limiting any definitive conclusions. The decreased eCPR success rates in the INCEPTION PP data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR. The probabilities of eCPR survival compared to cCPR using the per protocol data and incorporating the previously identified informative priors are shown in Table 3.

@@

This Bayesian analysis of the INCEPTION ITT data alone (analysis with a vague prior) suggests the presence of a non-negligible mortality probability benefit with eCPR, in the range of 73%, compared to cCPR. The probability that this improvement in survival exceeds 10% is almost 66%. The consideration of pertinent prior information from two previous comparable RCTs[@RN6751;@RN6759] results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%. Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability. In contrast to the INCEPTION original conclusion of similar survival effects between the two treatments, this re-analysis suggests only a modest probability of approximately 15% that survival probabilities between the two techniques are within a reasonably acceptable equivalency range. Additional uncertainty about the benefits or harms of eCPR do arise with the per protocol analysis, both with and without informative priors.

The INCEPTION researchers have addressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion. However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data, and prevents a full and comprehensive exploitation and updating of past knowledge. For trials such as INCEPTION that fail to meet statistical significance and are often incorrectly thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765]. Rather than dichotomizing results into statistical significance or not, with an obligatory loss of information and understanding, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.

Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to 
individual data from any of the trials limiting the possibility of examining specific subgroups. Also for the per protocol analysis, this data was only available for survival and not the primary INCEPTION outcome that included survival with intact neurological status. The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices. Given the existence of only 3 trials  estimating the very divergent baseline cCPR rates could not be reliably modeled. We choose not to perform a traditional Bayesian meta-analysis with
a semi-informative prior for the between-study variation  because our
goal was to illustrate how Bayesian principles could be informative
when applied to the analysis and interpretation of new trial data, in situations both with and without previous knowledge.

The "take home" message from this study is that standard statistical analyses resulting in a conclusion of "similar survival effects of eCPR to cCPR" may be overly simplified and potentially inaccurate. This Bayesian analysis demonstrates that at present definitive conclusions regarding the superiority, inferiority, or equivalence are impossible. Rather the possibility of a clinically meaningful benefit, or less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the residual uncertainties. The final dilemma confronting clinicians is that even Bayesian analyses of randomized trials provide only probability estimates for *average* treatment effects and not for the elusive *individual* treatment effect.

### 2023-02-13 manuscript/manuscript.tex
grammer corrections

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, enhanced, interior hidden]}{\end{tcolorbox}}\fi

@@

Three different informative priors were considered i) a N(0.32, 0.47)
ii) a N(0.43, 0.49) iii) a N(0.31, 0.46) distributions to represent all
the combined available RCT data\citep{RN6759, RN6751}, only the
ARREST\citep{RN6751} data and only the PRAGUE\citep{RN6759} data,
respectively

These different prior probabilities were updated using with the
INCEPTION ITT data to create the posterior distributions displayed in
Table 2. The probability for enhanced eCPR survival has increased to
80.4\% with a skeptical prior to 84.9\% with the enthusiastic prior and,
as expected, the associated uncertainty has been reduced, as reflected
by the narrower 95\% CrI, with the additional data. The posterior
probability of improved eCPR survival goes from 80\% with a skeptical
prior to 85\% with the enthusiastic prior.

The probability that the eCPR survival improvements exceed a minimum
10\% clinical threshold for improvement are 71\% and 77\% for the
skeptical and enthusiastic priors, respectively. The corresponding ROPE
probabilities are 16.8\% and 13.8\% (Table 2). Posterior probabilities
with the combined prior fell as expected between the results with the
skeptical and enthusiastic priors. The graphical presentations of these
results are shown in Figure 2.

INCEPTION did not report a per-protocol analysis. Such an analysis may
be helpful in assessing treatment efficacy, unfortunately at the risk of
an increased risk of bias by not respecting the ITT principle. From
INCEPTION Figure S4\citep{CPR2023a}, it appears that the per protocol
data for mortality is 13 survivors from 61 patients in the cCPR group
compared to 5 survivors from 46 patients receiving eCPR. With a vague
prior, the OR of increased survival with eCPR is decreased compared to
cCPR but with very wide CrI (OR 0.45, 95\% CrI 0.15 - 1.35), limiting
any definitive conclusions. The decreased eCPR success rates in the
INCEPTION PP data results in reduced posterior probabilities of eCPR
benefit and an increased probability of equivalence or benefit with
cCPR. The probabilities of eCPR survival compared to cCPR using the per
protocol data and incorporating the previous identified informative
priors are shown in Table 3.

@@

This Bayesian analysis of the INCEPTION ITT data alone (with a vague
prior) suggests the presence of a non-negligible probability of benefit
with eCPR, in the range of 73\%, compared to cCPR. The probability that
this improvement in survival exceeds 10\% is almost 66\%. The
consideration of pertinent prior information from two previous
RCTs\citep{RN6751, RN6759} in the field results in estimates of improved
eCPR survival probabilities, with intact neurological status, in the
vicinity of 80-85\%. Moreover depending on the choice of prior beliefs
included, the posterior probability of at least a 10\% improved eCPR
survival is in the range of 71 - 77\% probability. In contrast to the
INCEPTION original conclusion of similar survival effects between the
two treatments, this re-analysis suggests only a modest probability of
approximately 15\% that survival probabilities between the two
techniques are within a reasonably acceptable equivalency range.
Additional uncertainty about the benefits or harms of eCPR arise from
the per protocol analysis, both with and without informative priors.

The INCEPTION researchers have addressed an important clinical question
in the most challenging of research environments and are to be
congratulated on their trial design, its execution, and a nuanced
discussion. However the natural constraints of standard statistical
analyses, also known as null hypothesis significance testing, limits the
quantitative appreciation of their data, and prevents a full and
comprehensive exploitation and updating of past knowledge. For trials
such as INCEPTION that fail to meet statistical significance and are
often incorrectly thought of as ``negative'' trials, null hypothesis
significance testing favors confusion ``between absence of evidence and
evidence of absence''\citep{RN6765}. Rather than dichotomizing results
into statistical significance or not, with an obligatory loss of
information, Bayesian analysis concentrates on estimation of key outcome
differences with direct probability measures of their uncertainty.

Although this Bayesian analysis presents the strengths noted above, it
also had limitations. First, we did not have access to individual data
from any of the trials limiting the possibility of examining specific
subgroups. Also for the per protocol analysis, this data was only
available for survival and not the primary INCEPTION outcome that
included survival with intact neurological status. The threshold choices
for superiority and equivalence may be seen as arbitrary, although the
analyses can be easily repeated for different choices. Given the
existence of only 3 trials estimating the very divergent baseline cCPR
rates could not be reliably modeled. We choose not to perform a
traditional Bayesian meta-analysis with a semi-informative prior for the
between-study variation because our goal was to illustrate how Bayesian
principles could be informative when applied to the analysis and
interpretation of new trial data, both in the situations of no and
considerable previous knowledge.

The ``take home'' message from this study is that standard statistical
analyses resulting in a conclusion of ``similar survival effects'' may
be overly simplified and potentially inaccurate. This Bayesian analysis
demonstrates that any definitive conclusions regarding the superiority,
inferiority, or equivalence of eCPR to cCPR are lacking. Rather the
possibility of a clinically meaningful benefit, or less likely the
possibility of clinically meaningful harm, has not been reasonably
excluded and continued research is necessary to clarifying the remaining
uncertainties. The final dilemma confronting clinicians is that even
Bayesian analyses of randomized trials provide only probability
estimates for \emph{average} treatment effect and not for the elusive
\emph{individual} treatment effect.

>>>>

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, sharp corners, enhanced, frame hidden, breakable]}{\end{tcolorbox}}\fi

@@

Three different informative priors were considered i) a N(0.32, 0.47)
ii) a N(0.43, 0.49) iii) a N(0.31, 0.46) distributions to represent all
the combined available RCT data\citep{RN6759, RN6751}, only the
ARREST\citep{RN6751} data and only the PRAGUE\citep{RN6759} data,
respectively. These different prior probabilities were updated using
with the INCEPTION ITT data to create the posterior distributions
displayed in Table 2. The posteriorprobability for enhanced eCPR
survival has increased to 80.4\% with a skeptical prior, to 84.9\% with
the enthusiastic prior and, as expected, the associated uncertainty has
been reduced, as reflected by the narrower 95\% CrI, with the additional
data.

The probability that the eCPR survival improvements exceed a minimum
10\% clinical threshold for improvement are 71\% and 77\% for the
skeptical and enthusiastic priors, respectively. The corresponding ROPE
probabilities are 16.8\% and 13.8\% (Table 2). Posterior probabilities
with the combined prior were as expected between the results with the
skeptical and enthusiastic priors. The graphical presentations of these
results are shown in Figure 2.

INCEPTION did not report a per-protocol analysis. Such an analysis may
be helpful in assessing treatment efficacy, unfortunately at the risk of
an increased risk of bias by not respecting the ITT principle. From
INCEPTION Figure S4\citep{CPR2023a}, it appears that the per protocol
data for mortality is 13 survivors from 61 patients in the cCPR group
compared to 5 survivors from 46 patients receiving eCPR. With a vague
prior, the OR of increased survival with eCPR compared to cCPR is
decreased but with very wide CrI (OR 0.45, 95\% CrI 0.15 - 1.35),
limiting any definitive conclusions. The decreased eCPR success rates in
the INCEPTION PP data results in reduced posterior probabilities of eCPR
benefit and an increased probability of equivalence or benefit with
cCPR. The probabilities of eCPR survival compared to cCPR using the per
protocol data and incorporating the previously identified informative
priors are shown in Table 3.

@@

This Bayesian analysis of the INCEPTION ITT data alone (analysis with a
vague prior) suggests the presence of a non-negligible mortality
probability benefit with eCPR, in the range of 73\%, compared to cCPR.
The probability that this improvement in survival exceeds 10\% is almost
66\%. The consideration of pertinent prior information from two previous
comparable RCTs\citep{RN6751, RN6759} results in estimates of improved
eCPR survival probabilities, with intact neurological status, in the
vicinity of 80-85\%. Moreover depending on the choice of prior beliefs
included, the posterior probability of at least a 10\% improved eCPR
survival is in the range of 71 - 77\% probability. In contrast to the
INCEPTION original conclusion of similar survival effects between the
two treatments, this re-analysis suggests only a modest probability of
approximately 15\% that survival probabilities between the two
techniques are within a reasonably acceptable equivalency range.
Additional uncertainty about the benefits or harms of eCPR do arise with
the per protocol analysis, both with and without informative priors.

The INCEPTION researchers have addressed an important clinical question
in the most challenging of research environments and are to be
congratulated on their trial design, its execution, and a nuanced
discussion. However the natural constraints of standard statistical
analyses, also known as null hypothesis significance testing, limits the
quantitative appreciation of their data, and prevents a full and
comprehensive exploitation and updating of past knowledge. For trials
such as INCEPTION that fail to meet statistical significance and are
often incorrectly thought of as ``negative'' trials, null hypothesis
significance testing favors confusion ``between absence of evidence and
evidence of absence''\citep{RN6765}. Rather than dichotomizing results
into statistical significance or not, with an obligatory loss of
information and understanding, Bayesian analysis concentrates on
estimation of key outcome differences with direct probability measures
of their uncertainty.

Although this Bayesian analysis presents the strengths noted above, it
also had limitations. First, we did not have access to individual data
from any of the trials limiting the possibility of examining specific
subgroups. Also for the per protocol analysis, this data was only
available for survival and not the primary INCEPTION outcome that
included survival with intact neurological status. The threshold choices
for superiority and equivalence may be seen as arbitrary, although the
analyses can be easily repeated for different choices. Given the
existence of only 3 trials estimating the very divergent baseline cCPR
rates could not be reliably modeled. We choose not to perform a
traditional Bayesian meta-analysis with a semi-informative prior for the
between-study variation because our goal was to illustrate how Bayesian
principles could be informative when applied to the analysis and
interpretation of new trial data, in situations both with and without
previous knowledge.

The ``take home'' message from this study is that standard statistical
analyses resulting in a conclusion of ``similar survival effects of eCPR
to cCPR'' may be overly simplified and potentially inaccurate. This
Bayesian analysis demonstrates that at present definitive conclusions
regarding the superiority, inferiority, or equivalence are impossible.
Rather the possibility of a clinically meaningful benefit, or less
likely the possibility of clinically meaningful harm, has not been
reasonably excluded and continued research is necessary to clarifying
the residual uncertainties. The final dilemma confronting clinicians is
that even Bayesian analyses of randomized trials provide only
probability estimates for \emph{average} treatment effects and not for
the elusive \emph{individual} treatment effect.
